---
title: "Linear model and Lasso linear model simulation"
author: "Eliza Chai"
date: "06/09/2022"
output:
  pdf_document
---

```{r setup, include=F}
library(knitr)
knitr::opts_chunk$set(echo = T)

```

In this problem, you will use the package `simulator` to perform the following simulation study.

Let $X$ be a $n × p$ data matrix where each entry is generated from a normal distribution with mean 0 and standard deviation 1. Let $y$ be the vector of responses

$y = X\beta + \epsilon$,

with $\beta$ = (0,1,2,0,0) (and therefore $p = 5$) and $\epsilon$ a $n$-vector where each entry is generated from a normal distribution with mean 0 and standard deviation $\sigma$.

We decide to approach the estimation of $\beta$ with the following two methods:

-   Simple linear model (`lm`)

-   Lasso linear model: Suppose now you want to take advantage of the fact that some entries of the unknown vector $\beta$ are zero, i.e. $\beta$ is sparse. You therefore decide to apply a lasso linear model for the estimation of $\beta$. Given a data matrix `x` and an output vector `y`.

```{r, include=T, eval=F}
devtools::install_github("jacobbien/simulator")
library(glmnet)
cv.out <- cv.glmnet(x,y,alpha=1,nfolds=5) # Cross validated choice of the penalty
optimal_lambda = cv.out$lambda[which.min(cv.out$cvm)]
beta_est = as.numeric(glmnet(x,y,alpha=1, lambda = optimal_lambda)$beta) # Run model
beta_est #This is the lasso estimate of beta
```

For the two methods above, you want evaluate the estimation accuracy using the two metrics:

-   L2 norm: $\Sigma_{j}(\beta_j - \hat{\beta_j})^2$

-   Support recovery, i.e. the proportion of entries of $\beta$ that were correctly estimated to be equal to or different from zero (in other words, the proportion of entries $j = 1,...,p$ for which the following statement holds true: ($\beta_j > 0$ and $\hat{\beta_j} > 0$) or $(\beta_j = 0$ and $\hat{\beta_j} = 0$))

Run a simulation study with $n = 40$ and $\sigma^2= \{1,4,7,10\}$. Plot the results of the estimation accuracy (L2 norm and support recovery) in function of $\sigma^2= \{1,4,7,10\}$.

Comment on the results.

```{r, include=T}
library(purrr)
library(simulator)
library(glmnet)

make_mv_linear_model <- function(n, beta, sigma_sq) 
{
  new_model(          # Model constructor
    name = "mv_lm", 
    label = sprintf("n = %s, beta = %s, sigma_sq = %s", n, paste(beta, collapse = " "), sigma_sq),
    params = list(beta = beta, sigma_sq = sigma_sq, n = n),
    simulate = function(n, beta, sigma_sq, nsim)
    {
      sim_list =  map(1:nsim, function(ii){
        p = length(beta)
        x <- matrix(rnorm(n*p, mean = 0, sd = 1), nrow = n, ncol = p)
        y <- x %*% beta + rnorm(n, 0, sqrt(sigma_sq))
        list("x" = x, "y" = y)})
      
      return(sim_list) 
    }
  )
}


lse <- new_method("lse", "LSE  w/o intrcpt",
                  method = function(model, draw) {
                    yy <- draw$y
                    xx <- draw$x
                    fit <- lm(yy ~ xx - 1)
                    list(betahat = fit$coef)
                  })


lasso <- new_method("lasso", "LASSO  w/o intrcpt",
                    method = function(model, draw) {
                      yy <- draw$y
                      xx <- draw$x
                      
                      cv.out <- cv.glmnet(xx,yy,alpha=1,nfolds=5)   # Cross validated choice of the penalty
                      optimal_lambda = cv.out$lambda[which.min(cv.out$cvm)]
                      beta_est = as.numeric(glmnet(xx,yy,alpha=1, lambda = optimal_lambda)$beta)  # Run model 
                      
                      list(betahat = beta_est)
                    })

sq_err <- new_metric("l2", 
             "Squared error",
             metric = function(model, out) {
               sum((out$betahat - model$beta)^2)   # beta is a scalar
             })

prop_est <- new_metric("support", 
             "support recovery",
             metric = function(model, out) {
               p = length(model$beta)
               correct= 0
               for (i in 1:p){
                 if (model$beta[i] == 0 & out$betahat[i] == 0){
                   correct = correct + 1
                 }
                 else if (model$beta[i] > 0 & out$betahat[i] > 0){
                   correct = correct + 1
                 }
               }
               correct/p
             })


sim <- new_simulation("ls_vs_lasso", "l2 vs support") %>%
  generate_model(make_mv_linear_model,                       # Specify data generation function
                 n = 40,                                     # Data generation parameter 1
                 sigma_sq = as.list(c(1,4,7,10)),
                 beta = c(0,1,2,0,0),
                 vary_along = c("sigma_sq")) %>%
  simulate_from_model(nsim = 15)  %>%
  run_method(list(lse,lasso)) %>%
  evaluate(list(sq_err, prop_est))

sim %>% plot_eval(metric_name = "l2")
sim %>% plot_eval(metric_name = "support")

```

The lasso linear model without intercept has a smaller squared error compared to the linear model without intercept using the L2 norm matrices, meaning better estimation accuracy.

The lasso linear model without intercept has a higher support recovery compared to the linear model without intercept, meaning better estimation accuracy.

